---
title: "Stat 420 Final Project Report"
author: "Max Piazza, David Orona, Nithya Arumugam, Abhitej Bokka"
date: "2024-12-05"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Data Analysis Project: Modeling Used Car Prices


## Introduction

In an ever-volatile market where every dollar counts, the used car market represents a critical sector of the consumer industry. With rising consumer demand and an increasing variety of vehicles entering the secondary market, understanding the factors that influence used car prices is essential. Buyers seek to make informed decisions based on value for money, while sellers aim to maximize returns by accurately pricing their vehicles. Bridging this gap requires a data-driven approach to uncover the relationships between vehicle specifications and market pricing.

This project utilizes the “Vehicle dataset” by Nehal Birla, Nishant Verma, and Nikhil Kushwaha, available on Kaggle at: [https://www.kaggle.com/datasets/nehalbirla/vehicle-dataset-from-cardekho/data?select=Car+details+v3.csv](https://www.kaggle.com/datasets/nehalbirla/vehicle-dataset-from-cardekho/data?select=Car+details+v3.csv). 

This dataset aggregates detailed information on over 10,000 used cars, including key variables such as fuel type, transmission, engine capacity, mileage, and kilometers driven, alongside categorical variables like seller type, ownership history, and geographic location. 

Our analysis is driven by three core objectives:

* To model the relationship between vehicle specifications (e.g., mileage, transmission, and fuel type) and their pricing.
* To identify regional trends and seller-specific factors influencing market prices.
* To evaluate how performance metrics, such as engine power and fuel efficiency, impact purchasing behavior.

<!-- Once we know what exactly what "statistal modeling techinques" we use, we can refer to them properly in this introduction, for now I just have "regression analysis"  -->

Using statistical modeling techniques, including regression analysis, we aim to deliver a robust and interpretable model that not only predicts car prices but also highlights the most influential factors driving price variations. Our results will shed light on market dynamics, offering actionable insights for both consumers and industry professionals navigating this volatile space.

By the conclusion of this project, we aim to provide a detailed analysis that enhances understanding of the used car market, aiding stakeholders in making informed decisions in an ever-changing economic landscape.







```{r}

# Getting the proper libraries and reading data
library(tidyverse)
library(car)
library(broom)
library(MASS)
library(boot)
library(lmtest)

# I set up  all these packages beforehand
```



```{r, last}

library(broom)

# Step 1: Load and Prepare Data

car_details_v3 <- read_csv("dataset/Car details v3.csv")

# Data Cleaning and Transformation
car_details_v3_clean <- car_details_v3 %>%
  rename(
    Make = name,
    Year = year,
    Price = selling_price,
    Kilometer = km_driven,
    `Fuel Type` = fuel,
    `Seller Type` = seller_type,
    Transmission = transmission,
    Mileage = mileage,
    Engine = engine,
    `Max Power` = max_power,
    Seats = seats
  ) %>%
  drop_na(Price, Kilometer, Year, `Fuel Type`, Transmission) %>%
  mutate(
    Year = as.numeric(Year),
    Transmission = as.factor(Transmission),
    `Fuel Type` = as.factor(`Fuel Type`),
    `Seller Type` = as.factor(`Seller Type`)
  )

# Step 2: Create Mileage and Year Bins
car_details_v3_clean <- car_details_v3_clean %>%
  mutate(
    # Simplify Mileage: Convert to numeric and bin
    Mileage_numeric = as.numeric(gsub(" kmpl| km/kg", "", Mileage)),
    Mileage_bin = cut(Mileage_numeric, breaks = c(0, 10, 15, 20, 25, Inf), labels = c("0-10", "10-15", "15-20", "20-25", "25+")),
    # Bin Year into broader categories
    Year_bin = case_when(
      Year >= 1995 & Year <= 1999 ~ "1995-1999",
      Year >= 2000 & Year <= 2004 ~ "2000-2004",
      Year >= 2005 & Year <= 2009 ~ "2005-2009",
      Year >= 2010 & Year <= 2014 ~ "2010-2014",
      Year >= 2015 & Year <= 2020 ~ "2015-2020",
      TRUE ~ "Other"
    ),
    Year_bin = factor(Year_bin, levels = c("1995-1999", "2000-2004", "2005-2009", "2010-2014", "2015-2020", "Other"))
  )

# Step 3: Exploratory Data Analysis
# Obtaining Summary statistics
summary(car_details_v3_clean)

# Defining a function to remove outliers based on IQR
remove_outliers <- function(data, column) {
  Q1 <- quantile(data[[column]], 0.25, na.rm = TRUE)
  Q3 <- quantile(data[[column]], 0.75, na.rm = TRUE)
  IQR <- Q3 - Q1
  data %>%
    filter(data[[column]] >= (Q1 - 1.5 * IQR) & data[[column]] <= (Q3 + 1.5 * IQR))
}

# Removing outliers for plots
car_details_no_outliers <- car_details_v3_clean %>%
  remove_outliers("Price") %>%
  remove_outliers("Kilometer")

# Visualization: Kilometer vs Price without outliers
ggplot(car_details_no_outliers, aes(x = Kilometer, y = Price, color = `Fuel Type`)) +
  geom_point(alpha = 0.6) +
  labs(title = "Kilometer vs Price by Fuel Type (No Outliers)")

# Visualization: Distribution of Mileage Bins without outliers
ggplot(car_details_no_outliers, aes(x = Mileage_bin)) +
  geom_bar() +
  labs(title = "Distribution of Mileage Bins (No Outliers)") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

# Visualization: Boxplot of Price by Year Bins without outliers
ggplot(car_details_no_outliers, aes(x = Year_bin, y = Price)) +
  geom_boxplot() +
  labs(title = "Price Distribution by Year Bins (No Outliers)", x = "Year Bins", y = "Price")


# Step 4: Develop and Refine Model
# Our Initial model
model_initial <- lm(Price ~ Kilometer + `Fuel Type` + Transmission + Mileage_bin + Seats + Year_bin, data = car_details_v3_clean)

# Checking for multicollinearity using VIF
vif_initial <- vif(model_initial)
print(vif_initial)

# Refining model using stepwise selection
model_refined <- step(model_initial, direction = "both")
summary(model_refined)

# Step 5: Residual Diagnostics
# Diagnostic Plots
par(mfrow = c(2, 2))
plot(model_refined)
par(mfrow = c(1, 1))

# Testing for normality of residuals using Q-Q plot
qqnorm(residuals(model_refined))
qqline(residuals(model_refined), col = "red")

# Testing for homoscedasticity using Breusch-Pagan test
bp_test <- bptest(model_refined)
print(bp_test)

# Step 6: Model Validation
# Spliting data into training and testing sets
set.seed(123)
train_indices <- sample(1:nrow(car_details_v3_clean), size = 0.7 * nrow(car_details_v3_clean))
train_data <- car_details_v3_clean[train_indices, ]
test_data <- car_details_v3_clean[-train_indices, ]

# Training model
model_train <- lm(Price ~ Kilometer + `Fuel Type` + Transmission + Mileage_bin + Seats + Year_bin, data = train_data)

# Working with predictions
predictions <- predict(model_train, newdata = test_data)
actuals <- test_data$Price

# Calculating RMSE and R-squared
rmse <- sqrt(mean((predictions - actuals)^2))
r_squared <- 1 - (sum((predictions - actuals)^2) / sum((actuals - mean(actuals))^2))

cat("RMSE:", rmse, "\nR-squared:", r_squared, "\n")

# Step 7: Confidence Intervals and ANOVA
# Confidence intervals for coefficients
conf_intervals <- confint(model_refined)
print(conf_intervals)

# Obtaining ANOVA for variable significance
anova_results <- anova(model_refined)
print(anova_results)

# Step 8: Final Model Interpretation
cat("Final Model Summary:\n")
summary(model_refined)





```












## Methods


## Results


## Discussion


## Appendix

