---
title: "Stat 420 Final Project Proposal <br> Used Car Pricing"
output:
  html_document:
    theme: readable
  pdf_document: default
urlcolor: cyan
---

<style>
h1.title {
  text-align: center;
  margin-top: 50px;
}
div.author {
  text-align: center;
}
</style>

<div class="author">
Max Piazza <br> David Orona <br> Nithya Arumugam <br> Abhitej Bokka
</div>


## Introduction

This Vehicle dataset is available in kaggle [Vehicle Dataset from Cardekho- Cardetails Version3](https://www.kaggle.com/datasets/nehalbirla/vehicle-dataset-from-cardekho?select=Car+details+v3.csv).

The dataset contains the details of used cars.It includes various attributes like the car's make, manufacturing year, fuel efficiency, fuel type, transmission type, and more.

**Objective:**
The goal is to analyze the dataset to identify key factors influencing the price of a used car and build a predictive model for estimating its cost. 

The dataset contains the following attributes 

|    | Attributes       | Description                                                                        |
|----|------------------|------------------------------------------------------------------------------------|
| 1  |   Name           |   The brand name of the vehicle (e.g., Hyundai i10, Honda City).                   |
| 2  |   Year           |   The year the vehicle was manufactured.                                           |
| 3  |   Selling Price  |   The selling price of the vehicle.                                                |
| 4  |   Km Driven      |   The total distance the vehicle has been driven.                                  |
| 5  |   Fuel           |   The type of fuel used by the vehicle (e.g Petrol, Diesel)                        |
| 6  |   Seller Type    |   The type of seller (e.g Individual, Dealer).                                     |
| 7  |   Transmission   |   The type of transmission system (e.g Manual, Automatic).                         |
| 8  |   Owner          |   The number of previous owners of the vehicle (e.g., First Owner, Second Owner).  |
| 9  |   Mileage        |   The fuel efficiency of the vehicle.(in km/l or km/kg)                            |
| 10 |   Engine         |   The engine displacement capacity.(in CC)                                         |
| 11 |   Max Power      |   The maximum power output of the vehicle’s engine measured in horsepower.(in HP). |
| 12 |   Torque         |   The maximum torque.                                                              |
| 13 |   Seats          |   The seating capacity of the vehicle (e.g 5-seater, 7-seater).                    |
|    |                  |                                                                                    |

The above attributes can be categorized into Numeric and Categorical variables, with **selling price as continuous numeric response variable.**

| Attribute Type               | Attribute                                         |
|------------------------------|---------------------------------------------------|
| Categorical Variables        | Name, Fuel, Seller Type, Transmission, Owner      |
| Discrete Numeric Variables   | Year, Km Driven, Seats                            |
| Continuous Numeric Variables | Selling Price, Mileage, Engine, Max Power, Torque |

### Import dataset
```{r message=FALSE}
library(readr)
car_details = read_csv("Car details v3.csv")
```

### Total number of instances
```{r}
nrow(car_details)
```

### Excluding missing values from dataset
```{r eval=FALSE, include=FALSE}
col_name=""
for(col_name in names(car_details))
{
  if(sum(is.na(car_details[col_name]))>0)
  {
    print(c(col_name,":",sum(is.na(car_details[col_name]))))
  }
}
```

```{r echo=TRUE}
car_details = na.omit(car_details)
nrow(car_details)
```

### Excluding duplicate rows from dataset
```{r}
sum(duplicated(car_details))
car_details = car_details[!duplicated(car_details),]
nrow(car_details)
```
### Sample data from Vehicle's dataset
```{r}
head(car_details)