---
title: "Stat 420 Final Project Proposal"
author: "Max Piazza, David Orona, Nithya Arumugam, Abhitej Bokka"
date: "2024-11-26"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Data Analysis Project: Used Car Pricing

Our group will use the "Vehicle dataset" by Nehal Birla, Nishant Verma, Nikhil Kushwaha available on kaggle.com at: 

https://www.kaggle.com/datasets/nehalbirla/vehicle-dataset-from-cardekho/data?select=Car+details+v3.csv

## Description of the Data File

This project will utilize multiple datasets to ensure both comprehensiveness and robustness of analysis:

  1. car details v4.csv
  
  - Number of Variables: 20
  - Variables
    - Categorical
      - Make: Brand of the car
      - Model: Specific model of the car
      - Fuel type: Type of fuel (ex. gas, diesel, electric)
      - Transmission: Manual or automatic
      - Location: Geographic location where car is listed for sale
      - Color: Color of the car
      - Owner: Owner position indicator
      - Seller type: Type of seller listing the car
      - Drivetrain: FWD, RWD, or AWD
    - Continuous
      - Price: Selling price of the car
      - Kilometer: Distance the car has traveled
      - Engine power
      - Max power: Maximum power output of the engine
      - Max torque: Maximum torque generated by the engine
      - Length: Length of the car
      - Width: Width of the car
      - Height: Height of the car 
    - Discrete 
      - Year: Manufacturing year of the vehicle
      - Seating capacity: Number of seats in the car
      - Fuel tank capacity: Maximum fuel capacity of the vehicle
    - Numeric response
      - Price: Selling price of the car
  - Number of Records: Filtering out missing values, there are around 1875 instances.

<br>

  2. Car details v3.csv
  
  - Number of Variables: 13
  - Variables
    - Categorical
      - Name: Manufacturer and model
      - Fuel: Type of fuel (ex. gas, diesel, electric)
      - Seller_type: Type of seller listing the car
      - Transmission: Manual or automatic
      - Torque: Force generated by the engine
    - Continuous
      - Km_driven: Total kilometers the car has been driven
      - Mileage: Distance covered per unit of fuel
      - Engine: Engine capacity
      - Max_power: Maximum power output of the engine
    - Discrete
      - Year: Year of manufacture
      - Owner: Owner position indicator
      - Seats: Seating capacity of the car 
    - Numeric response
      - Selling_price: Price at which the car is being sold
  - Number of Records: Excluding missing values and duplicate rows, there are 6717 instances.
      
<br>        
        
  3. CAR DETAILS FROM CAR DEKHO.csv and car data.csv
  
  - These datasets provide additional records and variables that can supplement gaps in the primary datasets. 
  For example, CAR DETAILS FROM CAR DEKHO.csv contains 4,340 rows with key variables like "name", "selling_price", and "fuel".
  
## Statement of Interest

In an ever-violate market where every dollar counts, the used car market represents a critical sector of the consumer industry. Influenced by a wide variety of factors (see "Description of the Data File"), this project aims to answer:

- The relationship between vehicle specifications (ex. fuel type, mileage, transmission, etc.) and their pricing.
- Regional and seller-type trends in the market
- Performance metrics and their specific impact on purchasing behavior
  
## The Data
```{r}
library(readr)
car_details_v4 = read_csv("dataset/car details v4.csv")
car_details_v3 = read_csv("dataset/Car details v3.csv")
head(car_details_v4)
head(car_details_v3)
```

    