---
title: "Stat 420 Final Project Proposal <br> Used Car Pricing"
output: 
  html_document:
    theme: readable
  pdf_document: default
urlcolor: cyan
---

<style>
h1.title {
  text-align: center;
  margin-top: 50px;
}
div.author {
  text-align: center;
}
</style>

<div class="author">
Max Piazza <br> David Orona <br> Nithya Arumugam <br> Abhitej Bokka
</div>


## Introduction

This Vehicle dataset is available in kaggle [Vehicle Dataset from Cardekho- Cardetails Version3](https://www.kaggle.com/datasets/nehalbirla/vehicle-dataset-from-cardekho?select=Car+details+v3.csv).

The dataset contains the details of used cars.It includes various attributes like the car's make, manufacturing year, fuel efficiency, fuel type, transmission type, and more.

**Objective:**
The goal is to analyze the dataset to identify key factors influencing the price of a used car and build a predictive model for estimating its cost. 

The dataset contains the following attributes 

|    | Attributes       | Description                                                                        |
|----|------------------|------------------------------------------------------------------------------------|
| 1  |   Name           |   The brand name of the vehicle (e.g., Hyundai i10, Honda City).                   |
| 2  |   Year           |   The year the vehicle was manufactured.                                           |
| 3  |   Selling Price  |   The selling price of the vehicle in INR.                                               |
| 4  |   Km Driven      |   The total distance the vehicle has been driven.                                  |
| 5  |   Fuel           |   The type of fuel used by the vehicle (e.g Petrol, Diesel)                        |
| 6  |   Seller Type    |   The type of seller (e.g Individual, Dealer).                                     |
| 7  |   Transmission   |   The type of transmission system (e.g Manual, Automatic).                         |
| 8  |   Owner          |   The number of previous owners of the vehicle (e.g., First Owner, Second Owner).  |
| 9  |   Mileage        |   The fuel efficiency of the vehicle.(in km/l or km/kg)                            |
| 10 |   Engine         |   The engine displacement capacity.(in CC)                                         |
| 11 |   Max Power      |   The maximum power output of the vehicle’s engine measured in horsepower.(in HP). |
| 12 |   Torque         |   The maximum torque.                                                              |
| 13 |   Seats          |   The seating capacity of the vehicle (e.g 5-seater, 7-seater).                    |
|    |                  |                                                                                    |

The above attributes can be categorized into Numeric and Categorical variables, with **selling price as continuous numeric response variable.**

| Attribute Type               | Attribute                                         |
|------------------------------|---------------------------------------------------|
| Categorical Variables        | Name, Fuel, Seller Type, Transmission, Owner      |
| Discrete Numeric Variables   | Year, Km Driven, Seats                            |
| Continuous Numeric Variables | Selling Price, Mileage, Engine, Max Power, Torque |

### Import dataset
```{r message=FALSE}
library(readr)
car_details = read_csv("Car details v3.csv")
```

### Total number of instances
```{r}
nrow(car_details)
```

### Excluding missing values from dataset
```{r eval=FALSE, include=FALSE}
col_name=""
for(col_name in names(car_details))
{
  if(sum(is.na(car_details[col_name]))>0)
  {
    print(c(col_name,":",sum(is.na(car_details[col_name]))))
  }
}
```

```{r echo=TRUE}
car_details = na.omit(car_details)
nrow(car_details)
```

### Excluding duplicate rows from dataset
```{r}
sum(duplicated(car_details))
car_details = car_details[!duplicated(car_details),]
nrow(car_details)
```
### Sample data from Vehicle's dataset
```{r}
head(car_details)
```

### Data cleaning
```{r}
library(stringr)

# Extract the first word from "name" 
car_details$name = word(car_details$name,1)
# Rename name to make
colnames(car_details)[1] = "make"
# Change datatypes of "make" from character to factor
car_details$make = as.factor(car_details$make)

# Extract the numeric value from mileage
car_details$mileage = word(car_details$mileage,1)
#Change data_type of "mileage" from character to numeric
car_details$mileage=as.numeric(car_details$mileage)

# Extract the numeric value from engine
car_details$engine = word(car_details$engine,1)
#Change data_type of "engine" from character to numeric
car_details$engine=as.numeric(car_details$engine)

# Extract the numeric value from max_power
car_details$max_power = word(car_details$max_power,1)
#Change data_type of "max_power" from character to numeric
car_details$max_power=as.numeric(car_details$max_power)

# Change datatypes of "fuel", "seller_type", "transmission"	
# From character to factor
car_details$fuel=as.factor(car_details$fuel)
car_details$seller_type=as.factor(car_details$seller_type)
car_details$transmission=as.factor(car_details$transmission)

# Encode numeric values for 1st/2nd/.. owner and convert to factor data type
car_details$owner=as.factor(car_details$owner)

##KM driven in ten thousands
car_details$km_driven = car_details$km_driven/10000
#Rename  km_driven
colnames(car_details)[4]="km_driven_in_10k"

#Dropping column torque
car_details = car_details[,!names(car_details) %in% "torque"]

#Selling price in ten thousands
car_details$selling_price = car_details$selling_price/10000
#Rename Selling price
colnames(car_details)[3]="selling_price_in_10k"
```

### Sample data
```{r}
head(car_details)
```


### Data analysis
#### Number of cars in each make
```{r echo=FALSE}
make_counts = sort(table(car_details$make),decreasing = TRUE)
make_counts_df = data.frame(make = names(make_counts), count = as.numeric(make_counts))
make_counts_df
```

### Creating new varaible make_category

- To reduce the model complexity and to increase interpretability, the car make can be grouped into broader categories as "Budget", "Mid-Range" or "Luxury" depending on general market perception

```{r}
car_details$make_category = ifelse(car_details$make %in% 
                                     c("Maruti", "Hyundai", "Tata", "Mahindra", "Datsun", 
                                        "Renault", "Chevrolet", "Fiat", "Kia", "Isuzu", 
                                        "Force", "Daewoo", "Ambassador", "Ashok"), 
                                    "Budget",
                             ifelse(car_details$make %in% 
                                      c("Honda", "Ford", "Toyota", "Volkswagen", "Nissan", 
                                        "Skoda", "Jeep", "Mitsubishi", "MG", "Land"), 
                                    "Midrange", 
                                    "Luxury"))
car_details$make_category = as.factor(car_details$make_category)
levels(car_details$make_category)
```
### Final structure of car_details 
```{r}
structure_car_details = data.frame(
  Column = names(car_details),
  Type = unname(sapply(car_details, class))
)
```

### Structure of car_details
```{r}
print(structure_car_details)
```

```{r}
barplot(sort(table(car_details$make)),horiz=TRUE,las=1,
        xlab ="Number of cars",
        ylab = "Car Make",
        col=rainbow(length(unique(car_details$make))),
        cex.names = 0.5,main="Num.of cars in each make")
```


### Selling Price distribution
```{r}
hist(car_details$selling_price_in_10k,xlab="Selling Price (in Ten Thousands)",main="Selling Price distribution")
```

### Selling Price based on make
```{r}
boxplot(selling_price_in_10k ~ make, data = car_details,col=rainbow(length(unique(car_details$make))),
        las = 2,cex.axis = 0.7,               
        main = "Selling Price by Make",
        xlab = "Car Make",
        ylab = "Selling Price (in 10k)")
```

```{r echo=FALSE}
par(mfrow = c(2, 2))
fuel_type_car_count = table(car_details$fuel)
seller_type_car_count = table(car_details$seller_type)

barplot(sort(fuel_type_car_count),horiz=TRUE,las=1,cex.names = 0.9,
        main="Num.of cars in each fuel type")
barplot(sort(seller_type_car_count),horiz=TRUE,las=1,cex.names = 0.8,
        main="Num.of cars in each seller_type")
trans_type_car_count = table(car_details$transmission)
owner_type_car_count = table(car_details$owner)
barplot(sort(trans_type_car_count),horiz=TRUE,las=1,cex.names = 0.8,
        main="Num.of cars in each transmission type")
barplot(sort(owner_type_car_count),horiz=TRUE,las=1,cex.names = 0.6,
        main="Num.of cars in each owner type")
fuel_type_car_count
seller_type_car_count
trans_type_car_count
owner_type_car_count
```

```{r echo=FALSE}
par(mfrow = c(1, 2))
boxplot(selling_price_in_10k ~ fuel, data=car_details,cex.axis = 0.8,
     main="SellingPrice based on fuel type")
plot(selling_price_in_10k ~ seller_type, data=car_details,cex.axis = 0.6,
     main="SellingPrice based on seller type")
par(mfrow = c(1, 2))
plot(selling_price_in_10k ~ transmission, data=car_details,
     main="SellingPrice based on trans. type")
plot(selling_price_in_10k ~ owner, data=car_details,las=2,cex.axis = 0.5,
     main="SellingPrice based on owner type")
```



### Selling Price based on year with Transmission Type

```{r}
colours = ifelse(car_details$transmission == "Automatic", "blue", "red")
plot(selling_price_in_10k ~ year,data=car_details,col=colours,pch=19,
     main = "Selling Price by Year with Transmission Type",
     xlab = "Year",
     ylab = "Selling Price (in Ten Thousands)")
legend("topleft", legend = c("Automatic", "Manual"),
       col = c("blue", "red"), pch = 19)
```

### Selling Price based on year with Owner Type

```{r}
colours = ifelse(car_details$owner == "First Owner", "red",
          ifelse(car_details$owner == "Second Owner", "blue",
          ifelse(car_details$owner == "Third Owner", "green",
          ifelse(car_details$owner == "Fourth & Above Owner", "orange", 
          "yellow"))))
plot(selling_price_in_10k ~ year,data=car_details,col=colours,pch=19,
     main = "Selling Price by Year with Owner Categories",
     xlab = "Year",
     ylab = "Selling Price (in Ten Thousands)")

legend("topleft", legend = c("First Owner","Second Owner","Third Owner","Fourth & Above Owner","Test Drive Car"),
       col = c("red", "blue", "green", "orange", "yellow"), pch = 19)
```


```{r echo=FALSE}
par(mfrow = c(1, 2))
plot(selling_price_in_10k ~ km_driven_in_10k, data=car_details,
     main="Price based on km_driven")
plot(selling_price_in_10k ~ mileage, data=car_details,
     main="Price based on Mileage")

par(mfrow = c(1, 2))
plot(selling_price_in_10k ~ engine, data=car_details,
     main="Price based on engine")
plot(selling_price_in_10k ~ max_power, data=car_details,
     main="Price based on max_power")
```

### Correlation of the numeric variables in vehicle dataset
```{r}
pairs(selling_price_in_10k ~ year+km_driven_in_10k+mileage+engine+max_power+seats ,data=car_details)
cor_mat = cor(car_details[,sapply(car_details,is.numeric)])
cor_mat
high_correlation = cor_mat[cor_mat > 0.5 & cor_mat != 1]
high_cor_indices = which(cor_mat > 0.5 & cor_mat != 1, arr.ind = TRUE)
data.frame(
  row = rownames(cor_mat)[high_cor_indices[, 1]],
  column = colnames(cor_mat)[high_cor_indices[, 2]],
  correlation = high_correlation)
```

### Split data into train and test
```{r}
set.seed(125)  
train_indices = sample(nrow(car_details), size = 0.80 * nrow(car_details))
train_data = car_details[train_indices, ]
test_data = car_details[-train_indices, ]
nrow(train_data)
nrow(test_data)
```


```{r}
price_int_model = lm(selling_price_in_10k ~ make_category*year*max_power*transmission + owner + seller_type + km_driven_in_10k, data=train_data)
summary(price_int_model)
plot(price_int_model,which=1)
```

### BIC backward interaction model
```{r}
n= nrow(train_data)
bic_model_int = step(price_int_model,direction="backward",k=log(n),trace=0)
summary(bic_model_int)
plot(bic_model_int,which=1)
```

### Influential points1
```{r}
indicies_to_exclude = unname(which(cooks.distance(bic_model_int) > 4/length(cooks.distance(bic_model_int))))
sum(cooks.distance(bic_model_int) > 4/length(cooks.distance(bic_model_int)))
train_data_filtered = train_data[-indicies_to_exclude,]
```

### Refitting the model without influential points
```{r}
#price_int_model = lm(selling_price_in_10k ~ make_category*year*max_power*transmission + owner + seller_type + km_driven_in_10k, data=train_data_filtered)
#summary(price_int_model)
#plot(price_int_model,which=1)
```

### BIC backward interaction model(without infuential data)
```{r}
#n= nrow(train_data_filtered)
#bic_model_int = step(price_int_model,direction="backward",k=log(n),trace=0)
#summary(bic_model_int)
#plot(bic_model_int,which=1)
```


